---
title: "Survival on the RMS Titanic with arborist"
author: "Przemyslaw Biecek"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{Survival on the RMS Titanic}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

# Data for Titanic survival

Let's see an example for `DALEX` package for classification models for the survival problem for Titanic dataset.
Here we are using a dataset `titanic` available in the `DALEX` package. Note that this data was copied from the `stablelearner` package.

```{r}
library("DALEX")
head(titanic_imputed)
```

# Embeddings

Embeddings have gained enormous popularity in modelling with deep network models. It turns out that they can also be calculated for tree-based models!

Two types of embeddings are implemented in the `arborist` package and we will look at them below.

## Binary embeddings

```{r embeddings1}
library("arborist")

model_fr <- train(titanic_imputed, "survived", engine = "randomForest", ntree = 25)

embeddings_binary <- arborist::embed(titanic_imputed, model_fr, type = "binary")
str(embeddings_binary)
```

And let's see PCA transformation for these embeddings.

```{r embeddings2, message=FALSE, echo=FALSE}
library("ggplot2")

pca_embed <- prcomp(embeddings_binary)
df <- as.data.frame(pca_embed$x)
df$val <- titanic_imputed$survived
ggplot(df, aes(PC1, PC2, color = val)) + 
    geom_point() +
    DALEX::theme_ema()
```

## Shap embeddings

Currently we are using `treeshap` package and it does not handle correctly categorical variables.
So we use only numerical ones.

```{r embeddings3, message=FALSE}
small_titanic <- titanic_imputed[,c(2,5,6,7,8)]

model_fr <- train(small_titanic, "survived", engine = "randomForest", type = "regression")

embeddings_shap   <- arborist::embed(small_titanic, model_fr, type = "shap")
str(embeddings_shap$shaps)
```

And let's see PCA transformation for these embeddings.

```{r embeddings4, message=FALSE, echo=FALSE}
pca_embed <- prcomp(embeddings_shap$shaps)
df <- as.data.frame(pca_embed$x)
df$val <- titanic_imputed$survived
ggplot(df, aes(PC1, PC2, color = val)) + 
    geom_point() +
    DALEX::theme_ema()
```



# Simple model with arborist

Let's use the `arborist` to train the model.

```{r}
# prepare model
library("arborist")
model_fr <- train(titanic_imputed, "survived")
model_fr
```

Leaderboard

```{r}
model_fr$leaderboard
```

# Model performance

```{r}
model_performance(model_fr)
```

# Variable importance plots

Use the `variable_importance()` explainer to present the importance of particular features. Note that `type = "difference"` normalizes dropouts, and now they all start in 0.

```{r}
vi_rf <- model_parts(model_fr)
head(vi_rf)
plot(vi_rf)
```

# Variable effects

As we see the most important feature is `Sex`. Next three important features are `Pclass`, `Age` and `Fare`. Let's see the link between model response and these features.

Such univariate relation can be calculated with `variable_effect()`.

## Age

Kids 5 years old and younger have a much higher survival probability.

```{r}
vr_age  <- model_profile(model_fr, variables =  "age")
head(vr_age)
plot(vr_age)
```

## Passenger class

Passengers in the first-class have much higher survival probability.

```{r}
vr_class  <- model_profile(model_fr, variables =  "class")
plot(vr_class)
```

## Fare

Very cheap tickets are linked with lower chances.

```{r}
vr_fare  <- variable_profile(model_fr, variables =  "fare")
plot(vr_fare)
```

## Embarked

Passengers that embarked from C have the highest survival.

```{r}
vr_embarked  <- model_profile(model_fr, variables =  "embarked")
plot(vr_embarked)
```


# Instance level explanations

Let's see break-down explanation for model predictions for 8 years old male from 1st class that embarked from port C.

```{r}
new_passanger <- data.frame(
  class = factor("1st", levels = c("1st", "2nd", "3rd", "deck crew", "engineering crew", "restaurant staff", "victualling crew")),
  gender = factor("male", levels = c("female", "male")),
  age = 8,
  sibsp = 0,
  parch = 0,
  fare = 72,
  embarked = factor("Southampton", levels = c("Belfast", "Cherbourg", "Queenstown", "Southampton"))
)

sp_rf <- predict_parts(model_fr, new_passanger)
plot(sp_rf)
```

It looks like the most important feature for this passenger is `age` and `sex`. After all his odds for survival are higher than for the average passenger. Mainly because of the young age and despite being a male.


# Session info

```{r}
sessionInfo()
```
