---
title: "Value of players based on Fifa 22"
author: "Przemyslaw Biecek"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{Value of players based on Fifa 22}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

# Data from Fifa 22

Let's see an example for regression/pricing model.
Here we are using a dataset `fifa22` available in the `arborist` package. 

```{r}
library("arborist")
fifa <- fifa22[!duplicated(fifa22$short_name),]
rownames(fifa) <- fifa$short_name
fifa <- na.omit(fifa[,c(5,7,18:57)])
fifa$value_eur <- log10(fifa$value_eur)
attr(fifa, "na.action") <- NULL

head(fifa)
```

# Simple model with arborist

Let's use the `arborist` to train the model.

```{r}
# prepare model
# model_fifa <- train(fifa, "value_eur")
library(randomForest)
model_rf <- randomForest(value_eur ~ ., data = fifa, ntree = 50)
library(DALEX)
set.seed(1313)
fifa100 <- fifa[sample(1:nrow(fifa), 250),]
model_ex <- explain(model_rf, fifa100, fifa100$value_eur)
```

# Model performance

```{r}
model_performance(model_ex)
```

# Variable importance plots

Use the `variable_importance()` explainer to present the importance of particular features. Note that `type = "difference"` normalizes dropouts, and now they all start in 0.

```{r}
vi_rf <- model_parts(model_ex)
tail(vi_rf)
plot(vi_rf)
```

# Variable effects

As we see the most important feature is `age`. Next three important features are `skill_ball_control`, `movement_reactions`, `dribbling` and `defending`. Let's see the link between model response and these features.

Such univariate relation can be calculated with `variable_effect()`.

## Age

```{r}
vr_age  <- model_profile(model_ex, variables =  "age")
plot(vr_age)
```

## All others

```{r}
vr_age  <- model_profile(model_ex, 
                  variables =  c("age", "skill_ball_control",  "movement_reactions", 
                                 "dribbling", "defending"))
plot(vr_age)
```

# Instance level explanations

Let's see break-down explanation for model predictions for Robert Lewandowski.

```{r}
rl9 <- fifa["R. Lewandowski", ]

rl9_pp <- predict_parts(model_ex, rl9)
plot(rl9_pp)

rl9_sh <- predict_parts(model_ex, rl9, type = "shap")
plot(rl9_sh)

vr <- c("age", "skill_ball_control",  "movement_reactions", "dribbling", "defending")
rl9_age  <- predict_profile(model_ex, rl9, variables =  vr)
plot(rl9_age, variables =  vr)
```


# Session info

```{r}
sessionInfo()
```
